---
title: "DA5"
output: pdf_document
date: "2025-06-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
states = row.names(USArrests)
states
```
we got the names of rows - states of the US 

```{r}
names(USArrests)
```

here are the names of colums 

```{r}
apply(USArrests, 2, mean)
```

apply works the same as in python
2 means that we want this function to be applied for columns 

```{r}
apply(USArrests, 2, var)
```

the biggest variance is among assault column 

```{r}
arrestspca = prcomp(USArrests, scale = TRUE)
summary(arrestspca)
```
```{r}
names(arrestspca)
arrestspca$scale
arrestspca$center

arrestspca$rotation
```
A large absolute loading ( –0.583 for Assault on PC1) means that variable contributes heavily to that component.
- sing tells the direction 

We can see that 
PC1 is essentially an “overall crime level” axis (all four crimes load strongly and in the same direction).
PC2 contrasts UrbanPop (–0.873) against the other three (positive but smaller): a “rural vs urban” dimension.
PC3 is driven by Rape (0.818) versus the rest.
PC4 pits Assault (–0.743) against Murder (0.649).

```{r}
biplot(arrestspca, scale=0)
```
```{r}
arrestspca$rotation=-arrestspca$rotation
arrestspca$x=-arrestspca$x
biplot (arrestspca , scale =0)
```

```{r}
vari = arrestspca$sdev^2
vari

```

The variance

```{r}
pve = vari / sum(vari)
pve
```
62% of variance is explained by pc1, 25 by pc2, 9 by pc3 and only 4 by pc4

```{r}
plot(pve , xlab=" Principal Component ", ylab="Proportion of Variance Explained ", ylim=c(0,1), type='b')
```
this is illustrated in the graph. We again see the hockey stick. The elbow is probably PC3 because after it it started to be more horizontal 

```{r}
plot(cumsum(pve), xlab="Principal Component ", ylab="Cumulative Proportion of Variance Explained ", ylim=c(0,1), type='b')
```
this is for culminative one
