---
title: "Logistic Regression"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Logistic Regression

## Set Up

```{r pressure, echo=FALSE}
#import libraries
library(dplyr)
library(ggplot2)
library(repr) 
# Set the parameters for the plot sizes for all plots below
options(repr.plot.width=4, repr.plot.height=4)
```
```{r}
Pulse <- read.csv("~/IntroDataScience/R/D4/Data/Pulse.csv")
attach(Pulse)
```

## Exploring the Data
```{r}
summary(Pulse[,"Hgt"])
summary(Gender)
```
## Linear Model for predicting Gender based on Hgt
```{r}
gender.lm<-lm(Gender ~ Hgt)
summary(gender.lm)
```
```{r}
options(repr.plot.width=4, repr.plot.height=4)
plot(Gender~Hgt)
  abline(gender.lm)
```
```{r}
# To see more of the data points that fall on top of each other jitter them!
plot(Hgt, jitter(Gender, amount = 0.05))
  abline(gender.lm)
```
## Another Example Predicting Golf Putts Made
```{r}
PropMade=c(84/101,88/119,61/108,61/125,44/134)
PuttLength=c(3,4,5,6,7)
```
```{r}
odds=PropMade/(1-PropMade)
odds
4.941/0.488
```
## Logistic Regression Model
```{r}
table(Gender,Hgt)
```
```{r}
a <- table(Gender,Hgt)
b <- apply(a,2,sum) # find the column sums of the table
tab <- rbind(a,b) # add sums at the bottom of each column
tab
```
```{r}
gender.glm <- glm(Gender ~ Hgt, family = binomial)
summary(gender.glm)
```
```{r}
b0 <- coef(gender.glm)[1]
b1 <- coef(gender.glm)[2]
plot(Hgt, jitter(Gender, amount = 0.05))
  curve(exp(b0+b1*x)/(1+exp(b0+b1*x)),add=TRUE)
```

```{r}
#confidence interval for the betas
confint(gender.glm)
# exponentiation gives CI for the odds ratio
exp(confint(gender.glm))
```
## Multiple Logistic regression
Predict Survival in thte ICU based on Age, SysBP, and Pulse.

```{r}
detach(Pulse)
ICU <- read.csv("~/IntroDataScience/R/D4/Data/ICU.csv")
attach(ICU)
```
```{r}
model=glm(Survive~Age+SysBP+ICU$Pulse, family=binomial) 
# specify ICU$Pulse or R gets confused about which pulse variable
summary(model)
```
Examine the regressors one at a time.

```{r}
# Quick Boxplots
options(repr.plot.width=12, repr.plot.height=4)

par(mfrow=c(1,3)) # to combine all plots in one grid with c(nrows, ncols)
boxplot(Age~Survive, xlab="Survive", ylab="Age")
boxplot(SysBP~Survive, xlab="Survive", ylab="SysBP")
boxplot(ICU$Pulse~Survive, xlab="Survive", ylab="Pulse")
```
## Better Boxplots
```{r}
# Functionalize the plot we want
binaryboxplot <- function(data, xlab=NULL, ylab=NULL){
    # A fancier boxplot 
    # Inputs: A data frame data, 
    #         xlab to be used as group
    #         ylab to be used as regressor
    # Output: A boxplot with data points
    
    #data[xlab] makes a list 
    #so we need to extract just the values as a vector with unlist()  
    group <- unlist(data[xlab]) 
    regressor <- unlist(data[ylab])

    options(repr.plot.width=7, repr.plot.height=3)
    ggplot(data, aes(x=group, y=regressor, group=group)) + 
      geom_boxplot(outlier.colour="red") +
      geom_jitter(width=0.1) +
      coord_flip() +
      xlab(xlab) + ylab(ylab)
}

binaryboxplot(ICU, "Survive", "Age")
binaryboxplot(ICU, "Survive", "SysBP")
binaryboxplot(ICU, "Survive", "Pulse")    
```
## Model with Age only
```{r}
m1 <- glm(Survive~Age, family=binomial) 

options(repr.plot.width=4, repr.plot.height=4)
plot(jitter(Survive,amount=.03)~Age)
  curve(exp(m1$coefficient[1]+m1$coefficient[2]*x)/
        (1+exp(m1$coefficient[1]+m1$coefficient[2]*x)), add=TRUE)
```
## Model with SysBP only
```{r}
m2 <- glm(Survive~SysBP, family=binomial) 

options(repr.plot.width=4, repr.plot.height=4)
plot(jitter(Survive,amount=.03)~SysBP)
  curve(exp(m2$coefficient[1]+m2$coefficient[2]*x)/
        (1+exp(m2$coefficient[1]+m2$coefficient[2]*x)), add=TRUE)
```
## Model with Pulse only
```{r}
m3 <- glm(Survive~ICU$Pulse, family=binomial) 

options(repr.plot.width=4, repr.plot.height=4)
plot(jitter(Survive,amount=.03)~ICU$Pulse)
  curve(exp(m3$coefficient[1]+m3$coefficient[2]*x)/
        (1+exp(m3$coefficient[1]+m3$coefficient[2]*x)), add=TRUE)
```
## G-test for Overall Fit
```{r}
m1=glm(Survive~Age+SysBP+ICU$Pulse, family=binomial)
summary(m1)
anova(m1)
```
## Better G-test Code
Now we can get really fancy and write a function to do the G test
```{r}
# G - test function, so we only write the code for the test once
Gtest <- function(model) {
  # 
  # Inputs: Null Devance, Residual Deviance, Null df, Residual df
  # Outputs: G test statistic, DF, and P-value 
  # ToDO: Add test for to check if the model approproate for a G test!
 
  if((summary(model)$family[1] == "binomial")) {
    n <- length(anova(model)[,1])
    null.deviance <- anova(model)[1,4]    
    residual.deviance <- anova(model)[n,4]   
    null.df <- anova(model)[1,3]
    residual.df <- anova(model)[n,3]
    
    G <- null.deviance - residual.deviance
    df <- null.df - residual.df
    p <- pchisq(G, df, lower.tail=F)
    
    cat("G =", G, "DF =", df, "P-value =", p)
      
    } else {
  stop("This is not a logistic model!")  
  }  
}    

Gtest(m1)
m.bad <- glm(formula = Survive ~ Age + SysBP + ICU$Pulse, family = "poisson")
# try: Gtest(m.bad)
```
## Nested G test
Test the overall goodness of the model as we sequentially add predictors.
```{r}
m.full <- glm(Survive ~ SysBP + Sex + I(Sex*SysBP), family = "binomial")
summary(m.full)
Gtest(m.full)
```
```{r}
anova(m.full)
```
## Plot of fit for male vs female
```{r}
m2<-glm(Survive ~ SysBP + Sex + I(SysBP*Sex), family=binomial) 
options(repr.plot.width=8, repr.plot.height=8.5)
plot(Survive~SysBP,type="n") #null graph
points(jitter(Survive,amount=.03)[Sex==1]~SysBP[Sex==1],col="red") 
#female points are red
points(jitter(Survive,amount=.03)[Sex==0]~SysBP[Sex==0],col="blue") 
#male points are blue 
curve(exp(m2$coefficient[1]+m2$coefficient[2]*x)/
      (1+exp(m2$coefficient[1]+m2$coefficient[2]*x)),add=TRUE,col="blue")
curve(exp(m2$coefficient[1]+m2$coefficient[3]+
      (m2$coefficient[2]+m2$coefficient[4])*x)/
      (1+exp(m2$coefficient[1]+m2$coefficient[3]+
      (m2$coefficient[2]+m2$coefficient[4])*x)),add=TRUE,col="red")
```
## Plot of fit for infection vs no infection
```{r}
m3 <- glm(Survive ~ SysBP + Infection + I(SysBP*Infection), family=binomial) 
options(repr.plot.width=8, repr.plot.height=8.5)
plot(Survive~SysBP,type="n") #null graph
points(jitter(Survive,amount=.03)[Infection==1]~SysBP[Infection==1],col="red") 
#infection points are red
points(jitter(Survive,amount=.03)[Infection==0]~SysBP[Infection==0],col="blue") 
#non-infection points are blue 
curve(exp(m3$coefficient[1]+m3$coefficient[2]*x)/
      (1+exp(m3$coefficient[1]+m3$coefficient[2]*x)),add=TRUE,col="blue")
curve(exp(m3$coefficient[1]+m3$coefficient[3]+
      (m3$coefficient[2]+m3$coefficient[4])*x)/
      (1+exp(m3$coefficient[1]+m3$coefficient[3]+
      (m3$coefficient[2]+m3$coefficient[4])*x)),add=TRUE,col="red")
```
```{r}
# Create a new, binary variable from a continuous one
ICU1 <-ICU
ICU1$DangerBP <- 0
ICU1$DangerBP[ICU$SysBP>140] <- 1
ICU1
```

